---
title: "Basic Modelling"
author: "Jonathan Cordell"
date: "20/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tibble)
library(ggplot2)
library(knitr)
library(dplyr)
library(formatR)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

First I will set up the basic parameters of the model that are required
```{r parameters}
classSize <- 30
maxTime <- 100
#Rate of infection + background
ROI <- 0.01
BROI <- 0.01
#Numbers found in different talks for avg rate of moving from 1 category to next, then using exponential CDF to work out probability of moving from 1 category to the next in 1 time step
EtoP1StepProb <- pexp(1,rate=1/4.8)
PtoI1StepProb <- pexp(1,rate=1/0.68)
ItoR1StepProb <- pexp(1,rate=1/7)

```
Here I define the initial classroom
```{r Setup}
curTime <- 1
#S=Susceptible, E=Pre-Symptomatic, P=Infectious w/out Symptoms, I=Infectious w/ Symptoms, R=Recovered
class1 <- data.frame("S"=classSize,"E"=0,"P"=0,"I"=0,"R"=0)
```

Now I run the simulation for the next 99 timesteps, with people moving from one category to the next based on a binomial distribution, with n being based on the current timestep, and p being either defined above, or worked out as the force of infection in the case of S->I
```{r TimeSteps}
while (curTime<maxTime){
  #Lambda is the force of infection via closed system (i.e. infectiousness from other people in school) + some background prob of being infected from outside life
  lambda <- BROI + ROI/classSize*(class1[curTime,"P"] + class1[curTime,"I"])
  #Using binomial distribution to have random number of people moving to next stage via 
  StoE <- rbinom(1, class1[curTime,"S"], lambda)
  EtoP <- rbinom(1, class1[curTime,"E"], EtoP1StepProb)
  PtoI <- rbinom(1, class1[curTime,"P"], PtoI1StepProb)
  ItoR <- rbinom(1, class1[curTime,"I"], ItoR1StepProb)
  
  newS <- class1[curTime,"S"] - StoE
  newE <- class1[curTime,"E"] + StoE - EtoP
  newP <- class1[curTime,"P"] + EtoP - PtoI
  newI <- class1[curTime,"I"] + PtoI - ItoR
  newR <- class1[curTime,"R"] + ItoR
  
  class1 <- add_row(class1, "S"=newS, "E"=newE, "P"=newP, "I"=newI, "R"=newR)
  
  curTime=curTime+1
}
```
Plotting the values in each timestep as a stacked bargraph with a legend
```{r Plotting}
barplot(t(as.matrix(class1)), col=c("red","orange","yellow","green","cyan"), xlab="Days", ylab="Students")
legend("topright", colnames(class1), fill=c("red","orange","yellow","green","cyan"), bg="white")
```